---
title: "MOSAIC: sample size justification"
format: html
editor: visual
embed-resources: true
editor_options: 
  chunk_output_type: console
df-print: paged

---

## Sample Size Justification

In this parallel-group randomised trial, log-fold change (LFC) in peripheral blood VAF from baseline is the primary endpoint.

We anticipate that the placebo group will have on average no reduction in VAF (LFC = 0), compared with the intervention group with an average of 25% reduction in VAF (LFC = log (0.75)).

We assume that the LFC in VAF is approximately normally distributed. Under an exponential growth model : VAF(t)= c0 exp(rt), where t is time in years and r is the annual growth rate, the LFC at one-year post-baseline simplifies to log (VAF (1)/VAF (0)) = r. Based on Fabre et al. (2020), which characterised the longitudinal dynamics of CH, we estimated the within group LFC standard deviation to be 35%.

Under these assumptions, with a 2:1 randomisation and accounting for 10% attrition, 60 participants (40 Dec+THU, 20 placebo) provide at least 80% power to detect a 25% relative reduction in VAF using a two-sided significance of 0.05.

## Sample Size Estimation

Sample size calculation for a two-group comparison with unequal allocation.

$$ n_2 = \frac{(z_{\alpha/2} + z_{\beta})^2}{d^2} \cdot \left( 1 + \frac{1}{k} \right) $$ $$ n_1 = kn_2$$

```{r}
#| message: false
#| warning: false

library(tidyverse)

sample_size_fn1 <- function(alpha = 0.05, power = 0.8, ratio, ratio.sd, k = 2) {
  effect_size <- abs(log(ratio) / ratio.sd)

  z_alpha <- qnorm(1 - alpha / 2)

  z_beta <- qnorm(power)
  n2 <- ceiling((1 + 1 / k) * (z_alpha + z_beta)^2 / effect_size^2)
  n1 <- k * n2

  tibble(n1, n2)
}


```

```{r}
sz <- sample_size_fn1(power = 0.8, ratio = 0.75, ratio.sd = 0.35)
sz

```

Considering 10% attrition:

```{r}
sz/0.9

```

